---
title: "ST 558 - Fourth Blog Post"
author: "Maxwell Marion-Spencer"
date: '2022-07-19'
---

## Favorite Modeling Method: Boosted Tree

The modeling method I found the most interesting was the boosted tree method. 

This method relies on growing trees in a sequential manner, where each tree is based on an updated version of the previous tree, and the predictions are updated as the trees grow. A benefit of this method is that it is fairly slow, so you are unlikely to over fit the model.

There are four tuning parameters for a Boosted Tree model:

    n.trees: Number of trees, increasing n reduces the error on training set
    interaction.depth: number of splits it has to perform on a tree
    shrinkage: learning rate, reduce the impact of each additional fitted tree
    n.minobsinnode: minimum number of observations in trees’ terminal nodes


## A Boosted Tree Example:

First, we will bring our necessary libraries and the `iris` data set to use as our example data.

```{r, message=FALSE}
# Libraries
library(tidyverse)
library(caret)
library(gbm)

# Data set
data("iris")
```


Next, we will split our iris data into a test set and a training set. With 70% of our values going to our training set, and 30% going into our test set. 

```{r}
# Establishing train and test sets
set.seed(11)
train <- sample(1:nrow(iris),size=nrow(iris)*0.7)
test <- setdiff(1:nrow(iris),train)
irisTrain <- iris[train, ]
irisTest <- iris[test, ]
```


Typically we would conduct some exploratory data analysis on the training set, in order to get an idea of the distributions and existing relationships between the variables.

We can now fit our boosted tree model, with Sepal.Width, Petal.Length, and Petal.Width being used to predict Sepal.Length.

We are using five-fold cross validation.

```{r}
fit_boost <- train(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width,
                 data = irisTrain,
                 method = "gbm",
                 preProcess = c("center", "scale"),
                 trControl = trainControl(method = "cv", number = 5),
                 tuneGrid = expand.grid(n.trees = c(50, 100, 150),
                                        interaction.depth = 1:3, 
                                        shrinkage = 0.1, 
                                        n.minobsinnode = 10),
                 verbose = FALSE)

fit_boost
```


We can now apply our final model to our test set.

```{r}
prediction <- round(postResample(predict(fit_boost, newdata = irisTest), obs = irisTest$Sepal.Length),4)
prediction
```

This prediction seems pretty good! The RMSE is on the lower end, though we appear to have lost a bit of RSquared value compared to the performance on the training data. 

## Render Code

```{r, eval=FALSE, echo=FALSE}
# Render code I used to create this blog post
rmarkdown::render(
    '_Rmd/2022-07-19-blog-no4.Rmd', 
    output_format = "github_document",
    output_dir = "_posts",
    output_options = list(html_preview = FALSE)
)
```
